---
title: "Project: Analyzing World Happiness and GDP per Capita  "
author: ["Brian Mai", "Suhanth Alluri", "Zaíd Ruíz", "Maxwell Silver"]
format: 
  html:
    code-fold: true
toc: true
embed-resources: true
code-tools: true
execute:
  echo: true
  message: false
  warning: false
  error: true
---

## The Data

Data is taken from the following sources:

-   [Gapminder](https://www.gapminder.org/data/documentation/gd001/)
-   [World Happiness Report](http://gapm.io/dhapiscore_whr)

### Load appropriate libraries

```{r}
library(tidyverse)
library(knitr)
```

### Load Happiness Score Data

This dataset provides happiness score data for several countries across the years 2005 to 2023.

The happiness score is determined from the national average response to the question of life evaluations asking the following “Please imagine a ladder, with steps numbered from 0 at the bottom to 10 at the top. The top of the ladder represents the best possible life for you and the bottom of the ladder represents the worst possible life for you. On which step of the ladder would you say you personally feel you stand at this time?” This measure is also referred to as Cantril life ladder. Gapminder has converted this indicator's scale from 0 to 100 to easily communicate it in terms of percentage.

More information of the data can be found [here](https://docs.google.com/spreadsheets/d/1Hh2E8PrDA9nzRY_ikBAF3MSKkHreEUVLXnjE3PtJVVA/edit?gid=501532268#gid=501532268)

```{r}
happiness_score <- read_csv("hapiscore_whr.csv")
```

### Load GDP Per Capita data

This dataset provides GDP per capita data for several countries across the years 1800 to 2100. The future years' values are projceted GDP per capita values.

GDP per capita measures the value of everything produced in a country during a year, divided by the number of people. The unit is constant dollars adjusted for inflation in 2017’s prices. As the cost of living varies across countries, we use a currency called “international dollars”, which is adjusted for Purchasing Power Parity (PPP). This is a virtual currency that enables better comparisons. Such a dollar would buy in each country a comparable amount of goods and services as a U.S. dollar would buy in the United States. GDP per capita is the GDP divided by the population of the country, which gives a rough estimate of the average annual income of the citizens.

More information of the data can be found [here](https://www.gapminder.org/data/documentation/gd001/)

```{r}
gdp_per_capita <- read_csv("gdp_pcap.csv")
```


## 1.1 Data Cleaning

We first started off by cleaning the happiness data set. We saw that the 2005 column was the only row with more than 80% of its values equal to NA.

```{r}
happiness_score |>
  summarise(across(everything(), ~ mean(is.na(.)) > 0.80)) |>
  pivot_longer(
    cols = everything(),
    names_to = "column",
    values_to = "NA_more_80_percent") |>
  kable()
```

To clean the data, we simply removed the 2005 column as it does not add many observations to our data.
```{r}
happiness_score_clean <- happiness_score |>
  select(country, `2006`:`2023`)
```

We noticed that in the GDP per capita data some years had character values since their values used 'k' to denote thousands  (e.g., '24.5k' for 24,500). We cleaned this dataset by converting all such values to their exact numeric form, ensuring all data is now in a numeric data type. This way we are able to do computations with the GDP per capita data more easily.
```{r}
gdp_per_capita_clean <- gdp_per_capita |>
  mutate(across(`1800`:`2100`, ~ as.numeric(str_replace_all(., "k", "e3"))))
```

While some rows in both datasets contain NA values, we have decided to retain them in our clean datasets. If one variable is missing but the other is present, we choose to keep the row to allow us to analyze the trend of the variable with available data. If both variables are missing, we still retain the row to clearly indicate that the values are missing for that particular year.

## 1.2 Pivot Longer
Next, we pivoted the data longer so that each country had a row for each year and variable value.
```{r}
happiness_score_clean <- happiness_score_clean |>
  pivot_longer(
    cols = `2006`:`2023`,
    names_to = "year",
    values_to = "happiness_score")

gdp_per_capita_clean <- gdp_per_capita_clean |>
  pivot_longer(
    cols = `1800`:`2100`,
    names_to = "year",
    values_to = "gdp_per_capita")
```


## 1.3 Joining Datasets
Lastly, we joined the pivoted datasets.
```{r}
merged_data <- left_join(happiness_score_clean, gdp_per_capita_clean, by = c("country", "year"))

head(merged_data) |>
  kable()
```

## Data Relationship
We picked to analyze the relationship between happiness and GDP per capita because we were curious about the age old question: "Does money equal happiness?" 

### Hypothesis
According to [Forbes](https://www.forbes.com/sites/johnjennings/2024/02/12/money-buys-happiness-after-all/), it seems like people's happiness, which seemed to increase up until a $75,000 salary is reached, may actually still be improved even after reaching an income above this perceieved threshold, thus proving that someone's happiness is strongly infuenced by their income.  

Our hypothesis is that there is a strong positive correlation between happiness and GDP per capita. We expect that as GDP per capita increases, happiness score will also increase.
